---
title: "Asymmetric Poisson Pseudo-Maximum Likelihood (APPML)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Asymmetric Poisson Pseudo-Maximum Likelihood (APPML)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(capybara)
```

## Overview

Standard Poisson Pseudo-Maximum Likelihood (PPML) estimates the conditional mean of the outcome
variable. `fepoisson_asymmetric()` extends this by fitting conditional expectiles — a
generalization that lets you examine different parts of the conditional distribution with the
same efficiency gains of high-dimensional fixed effects.

The implementation follows Bergstrand et al. (2025): instead of minimizing a symmetric loss, the
estimator minimizes an *asymmetric* weighted loss controlled by a single parameter `expectile`
($\tau \in (0, 1)$):

- $\tau = 0.5$ — standard PPML (symmetric, estimates the conditional mean).
- $\tau < 0.5$ — places more weight on observations with *negative* residuals, pulling estimates
  toward the lower part of the distribution.
- $\tau > 0.5$ — places more weight on observations with *positive* residuals, shifting estimates
  toward the upper part of the distribution.

The `expectile` is set through `fit_control()` and defaults to `0.5`, so if you don't specify it,
`fepoisson_asymmetric()` will behave like standard PPML.

## Estimating lower and upper expectiles

Shifting `expectile` away from 0.5 lets you trace how the entire conditional distribution
of `mpg` varies with `wt`, after absorbing the `cyl` fixed effects.

```{r lower}
# 10th expectile — sensitive to small values of mpg
mod_low <- fepoisson_asymmetric(
  mpg ~ wt | cyl,
  data    = mtcars,
  control = fit_control(expectile = 0.1)
)
```

```{r ppml}
# Standard PPML: expectile = 0.5 recovers the conditional mean
mod_ppml <- fepoisson_asymmetric(
  mpg ~ wt | cyl,
  data    = mtcars,
  control = fit_control(expectile = 0.5)
)
summary(mod_ppml)
```

```{r upper}
# 90th expectile — sensitive to large values of mpg
mod_high <- fepoisson_asymmetric(
  mpg ~ wt | cyl,
  data    = mtcars,
  control = fit_control(expectile = 0.9)
)
```

## Comparing coefficients across expectiles

A natural use-case is comparing how the effect of a regressor changes across the distribution.
The table below collects the coefficient on `wt` at the three expectile levels:

```{r compare}
summary_table(
  mod_low, mod_ppml, mod_high,
  model_names = c("10th expectile", "PPML (50th expectile)", "90th expectile")
)
```

A negative coefficient that grows in magnitude from `low` to `high` would indicate that
heavier cars are disproportionately associated with low mpg at the top of the distribution —
a pattern that standard PPML would average out.

## Convergence diagnostics

The outer APPML loop updates observation weights until the coefficient vector stops changing.
You can inspect convergence through the returned list elements:

```{r diag}
cat("Outer loop converged:", mod_ppml$conv_outer, "\n")
cat("Outer iterations:    ", mod_ppml$iter_outer, "\n")
cat("Expectile used:      ", mod_ppml$expectile,  "\n")
```

## References

Bergstrand, Jeffrey H., Matthew W. Clance, and JMC Santos Silva. "The tails of gravity: Using
expectiles to quantify the trade-margins effects of economic integration agreements."
*Journal of International Economics* (2025): 104145.
