---
title: "Poisson Pseudo-Maximum Likelihood (PPML) model with cluster-robust standard errors"
output: rmarkdown::html_vignette
bibliography: "references.bib"
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We will estimate a Poisson Pseudo-Maximum Likelihood (PPML) model using the
data available in this package with the idea of replicating the PPML results
from Table 3 in @yotov2016advanced.

This requires to include exporter-time and importer-time fixed effects, and to
cluster the standard errors by exporter-importer pairs.

The PPML especification corresponds to:
\begin{align}
X_{ij,t} =& \:\exp\left[\beta_1 \log(DIST)_{i,j} + \beta_2 CNTG_{i,j} +\right.\\
\text{ }& \:\left.\beta_3 LANG_{i,j} + \beta_4 CLNY_{i,j} + \pi_{i,t} + \chi_{i,t}\right] \times \varepsilon_{ij,t}.
\end{align}

We use `dplyr` to obtain the log of the distance. This model excludes
domestic flows, therefore we need to subset the data also with `dplyr`.

Required packages:

```{r}
library(capybara)
library(dplyr)
```

Data wrangling:

```{r}
dppml <- trade_panel %>%
  mutate(
    log_dist = log(dist),
    exporter = substr(exp_year, 1, 3),
    importer = substr(imp_year, 1, 3)
  ) %>%
  filter(importer != exporter)
```

We can use the `fepoisson()` function to obtain the estimated coefficients
and we add the fixed effects as `| exp_year + imp_year` in the formula.

Model estimation:

```{r}
summary(fepoisson(
  trade ~ log_dist + cntg + lang + clny | exp_year + imp_year,
  data = dppml
))
```

The coefficients are almost identical to those in Table 3 from
@yotov2016advanced that were obtained with Stata. The difference is attributed
to the different fitting algorithms used by the two software.

To cluster the standard errors, we need to crete a pair variable and modify
the model formula to add a cluster term.

```{r}
dppml <- dppml %>%
  mutate(pair = paste(exporter, importer, sep = "-"))

summary(
  fepoisson(
    trade ~ log_dist + cntg + lang + clny | exp_year + imp_year | pair,
    data = dppml
  ),
  type = "cluster"
)
```

The result is similar and the difference is explained by the different
algorithms used to estimate the variance-covariance matrix. Capybara clustering
algorithm is based on @cameron2011robust while Stata uses a sandwich-type 
estimator.

# References
