---
title: "Different Variance-Covariance Estimators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Different Variance-Covariance Estimators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(capybara)
```

A very quick verification of Ross (2004) is to obtain the coefficients for the OLS model from table 1:

```{r verification1}
felm(ltrade ~ bothin + onein + gsp + ldist + lrgdp + lrgdppc + regional +
  custrict + comlang + border + landl + island + lareap + comcol +
  curcol + colony + comctry | year, data = ross2004)
```

This does not involve many fixed effects. However, capybara will be particularly useful to
obtain different standard errors for the same functional form, which is the main focus of
table 4B in Cameron & Miller (2014).

Table 4B shows the following clustering methods that we can replicate with the third part of the
model formula (e.g. `y ~ x1 + x2 + ... | fe1 + fe2 | cl1 + cl2`)  and using the `vcov` argument
to select how the clustering variables are used:

| `vcov` value     | Description                                    |
|:-----------------|:-----------------------------------------------|
| `"iid"`          | Default OLS (i.i.d. errors)                    |
| `"hetero"`       | Heteroskedastic-robust (HC0)                   |
| `"cluster"`      | One-way cluster sandwich                       |
| `"m-estimator"`  | One-way M-estimator sandwich                   |
| `"dyadic"`       | Dyadic-robust (Cameron & Miller, 2014)         |

Capybara provides an `update()` method to easily modify the formula for each model, as we need
to change the clustering variables for each model, which avoids error-prone copy-pasting of the full
formula every time.

## IID  (no cluster part in formula)

```{r table4b_models, cache = TRUE}
fml <- ltrade ~ bothin + onein + gsp + ldist + lrgdp + lrgdppc + regional +
  custrict + comlang + border + landl + island + lareap + comcol +
  curcol + colony + comctry | year

fit_iid <- felm(
  fml,
  data = ross2004, vcov = "iid"
)
```

## Heteroskedastic-robust (HC0)

```{r table4b_models2, cache = TRUE}
fit_hetero <- felm(
  fml,
  data = ross2004, vcov = "hetero"
)
```

## One-way: cluster on country-pair (g,h)

Note that multi-part formulas (e.g., `y ~ x | fe | cl`) are handled by the Formula package. Therefore,
updating those needs an explicit `Formula.as.Formula()` is the formula type was not stated.

```{r table4b_models3, cache = TRUE}
fml2 <- update(Formula::as.Formula(fml), . ~ . | pair)

fit_pairs <- felm(
  fml2,
  data = ross2004, vcov = "cluster"
)
```

## One-way: cluster on country 1 (g)

```{r table4b_models4, cache = TRUE}
fit_ctry1 <- felm(
  update(fml2, . ~ . | . | ctry1),
  data = ross2004, vcov = "cluster"
)
```

## One-way: cluster on country 2 (h)

```{r table4b_models5, cache = TRUE}
fit_ctry2 <- felm(
  update(fml2, . ~ . | . | ctry2),
  data = ross2004, vcov = "cluster"
)
```

## Two-way: cluster on (g) and (h) simultaneously

```{r table4b_models6, cache = TRUE}
fit_2way <- felm(
  update(fml2, . ~ . | . | ctry1 + ctry2),
  data = ross2004, vcov = "cluster"
)
```

## Dyadic-robust: Cameron-Miller (2014) sandwich with cross-dyad correlations

```{r table4b_models7, cache = TRUE}
fit_dyadic <- felm(
  update(fml2, . ~ . | . | ctry1 + ctry2),
  data = ross2004, vcov = "dyadic"
)
```

With the above models, we can replicate Table 4B from Cameron & Miller (2014) using
`summary_table()`, another convenience function in capybara to display multiple models side-by-side.
This completely avoids calling `summary()` on each model and then conducting post-processing to
extract the relevant information for the table.

```{r table4b_display}
summary_table(
  fit_iid, fit_hetero, fit_pairs, fit_ctry1, fit_ctry2, fit_2way, fit_dyadic,
  model_names = c("IID", "Hetero", "Pairs", "Country 1", "Country 2", "2-Way", "Dyadic")
)
```
